---
title: "Reproducible Research Week 2 Project"
author: "Daniel Malson"
date: "August 16th, 2020"
output: html_document
---

## Loading and preprocessing the data

```{r}
echo = TRUE 
## Loading and preprocessing the data
setwd("C:\\data")
activity_data <- read.csv("activity.csv", header=TRUE)
```

## What is mean total number of steps taken per day?

```{r}
## What is mean total number of steps taken per day?
echo = TRUE 
summary_data <- tapply(activity_data$steps, activity_data$date, sum, na.rm=TRUE)
hist(summary_data, xlab = "Steps Per Day", main = "Histogram of Summarized Activity Data")
```

The mean is:

```{r}
echo = TRUE 
mean_summary_data <- mean(summary_data, trim = 0) 
mean_summary_data <- round(mean_summary_data, digits = 2)
mean_summary_data
```

The median is:

```{r}
echo = TRUE 
median_summary_data <- median(summary_data, trim = 0) 
median_summary_data <- round(median_summary_data, digits = 2)
median_summary_data
```

## What is the average daily activity pattern?

```{r}
## What is the average daily activity pattern?

echo = TRUE
steps_interval_data <- aggregate(steps ~ interval, activity_data, mean)
plot(steps_interval_data$interval,steps_interval_data$steps, type="l", xlab = "Interval", ylab = "Steps", col= "blue")
```

## Imputing missing values

The number of records with missing steps:

```{r}
## Imputing missing values

echo = TRUE 
na_activity_data <- !complete.cases(activity_data)
sum_na_activity_data <- sum(na_activity_data)
sum_na_activity_data
```

Imput the missing values based interval:

```{r}

echo = TRUE 
loop_count <- 1
num_records <- nrow(activity_data)
num_records

no_na_activity_data <- activity_data


while (loop_count <= num_records) {
  
 steps <- activity_data$steps[loop_count]

 if(is.na(steps)) {

   interval_mean <- subset(steps_interval_data, interval == activity_data$interval[loop_count])  
   no_na_activity_data$steps[loop_count] <- round(interval_mean$steps)
 }
   
 loop_count <- loop_count +1
 
}
```

Verify that there are no missing values in the new data set: 

```{r}
## This is to verify there are no missing values for steps

echo = TRUE 
na_activity_data2 <- !complete.cases(no_na_activity_data)
sum_na_activity_data2 <- sum(na_activity_data2)
sum_na_activity_data2

```




```{r}
## Are there differences in activity patterns between weekdays and weekends?

echo = TRUE 

no_na_summary_data <- tapply(no_na_activity_data$steps, no_na_activity_data$date, sum)
hist(no_na_summary_data, xlab = "Steps Per Day", main = "Histogram of Summarized Activity Data (no missing steps values)")

# Mean with no missing steps values
no_na_mean_summary_data <- mean(no_na_summary_data, trim = 0) 
no_na_mean_summary_data <- round(no_na_mean_summary_data, digits = 2)
no_na_mean_summary_data

# Median with no missing steps values
no_na_median_summary_data <- median(no_na_summary_data, trim = 0) 
no_na_median_summary_data <- round(median_summary_data, digits = 2)
no_na_median_summary_data
```

Median difference:

```{r}
echo = TRUE 

# Median Diff
median_diff <- no_na_median_summary_data - median_summary_data
median_diff
```


Mean difference:

```{r}

# Mean Diff
mean_diff <- no_na_mean_summary_data - mean_summary_data
mean_diff
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
              
df_no_na_activity_data <- data.frame(no_na_activity_data)

df_no_na_activity_data$day_of_week = as.factor(ifelse(is.element(weekdays(as.Date(df_no_na_activity_data $date)),weekdays), "Weekday", "Weekend"))

agg_df_no_na_activity_data <- aggregate(steps ~ interval + day_of_week, df_no_na_activity_data, mean)

library(lattice)
xyplot(agg_df_no_na_activity_data$steps ~ agg_df_no_na_activity_data$interval|agg_df_no_na_activity_data$day_of_week, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```

